Unsupervised Learning for Identifying Events in Active Target Experiments

### 2020-06-26 doc/Manuscript/article.tex
Updates from Overleaf

\subsection{Classifying  events} 
\todo{Brute force serving as motivation for why we do unsupervised learning.}
%The filtering task relied on a set of $\chi^2$ distributions based on different criteria, 
The traditional event selection process involved a Monte-Carlo $\chi^2$ minimization method, however there was no well-defined method to quantify the effectiveness of the event selection from the thresholds applied to these distributions.
{\color{blue} Needs more explanation earlier in paper -MPK}
In addition, the selection task produced a binary result only, ``good" or``bad", relative to the event of interest, while the data clearly contained more classes of events%, if only because of the presence of carbon in the target (from the isobutane gas).
In a broader perspective, an unsupervised classification algorithm would offer the possibility to ``discover" rare events not anticipated (or overlooked) by the physicist. These events would likely be filtered out using the traditional methods. From a practical point of view, compared to supervised learning, it also avoids the necessary labeling task of the learning set events, which is error prone and time consuming.

@@

%We begin by investigating the K-means clustering algorithm on the VGG16 latent space. 
The VGG16 model is pre-trained on the ImageNet dataset, and subsequently applied to the AT-TPC events without the classification layers. This creates a set of vectors $\boldsymbol{x} \in \R^{8192}$ as our new representation of the events. To cluster we use \lstinline{scikit-learn} implementation of the K-means algorithm, with default parameters \cite{Pedregosa2011}.

@@

The MIXAE model comprises several autoencoders, each of which corresponds to a cluster. The number of autoencoders, and thus the number of clusters, has to be determined beforehand. Each autoencoder constructs a latent representation of a given example. Those representations are used as inputs to an auxiliary network which assigns scores to the clusters, indicating the likelihood that the given example belongs to each cluster. Examples are then assigned to the cluster with the highest score. The architecture is portrayed in Figure \ref{fig:mixae}, wherein tapered boxes denote a direction of compression in the network components.

@@

\noindent The objective during training is to enforce that these assumptions are met. For a more formal consideration of these heuristics, the model objective aand the assumptions made on the data by this model see \citet{Zhang}.

@@

\section{Results and Discussions}\label{sec:results}

@@

The same caveat is present in the MIXAE implementation. While it showed better optimal performance than the K-means method, some inconsistencies were observed, that were notably not evident from the unsupervised training-objectives of the model. These two factors currently conspire to limit its immediate applicability, and more developments are needed for this approach.

>>>>

\subsection{Classifying  events} 
\todo{Brute force serving as motivation for why we do unsupervised learning.}
%The filtering task relied on a set of $\chi^2$ distributions based on different criteria, 
The traditional event selection process involved a Monte-Carlo $\chi^2$ minimization method, however there was no well-defined method to quantify the effectiveness of the event selection from the thresholds applied to these distributions.
{\color{blue} Needs more explanation earlier in paper -MPK}
In addition, the selection task produced a binary result only, ``good" or``bad", relative to the event of interest, while the data clearly contained more classes of events.%, if only because of the presence of carbon in the target (from the isobutane gas).
In a broader perspective, an unsupervised classification algorithm would offer the possibility to ``discover" rare events not anticipated (or overlooked) by the physicist. These events would likely be filtered out using the traditional methods. From a practical point of view, compared to supervised learning, it also avoids the necessary labeling task of the learning set events, which is error prone and time consuming.

@@

%We begin by investigating the K-means clustering algorithm on the VGG16 latent space. 
The user a VGG16 model that is pre-trained on the ImageNet dataset and subsequently applied to the AT-TPC events without the classification layers. This creates a set of vectors $\boldsymbol{x} \in \R^{8192}$ as our new representation of the events. To cluster, we use the \lstinline{scikit-learn} implementation of the K-means++ algorithm with default parameters \cite{Pedregosa2011}.

@@

The MIXAE model comprises several autoencoders, each of which corresponds to a cluster. The number of autoencoders, and thus the number of clusters, has to be determined beforehand. Each autoencoder constructs a latent representation of a given example. Those representations are used as inputs to an auxiliary network which assigns scores to the clusters, indicating the likelihood that the given example belongs to each cluster. Examples are then assigned to the cluster with the highest score.

@@

\noindent The objective during training is to enforce that these assumptions are met. The architecture is portrayed in Figure \ref{fig:mixae}, wherein tapered boxes denote a direction of compression in the network components.

For a more formal consideration of these heuristics, the model objective aand the assumptions made on the data by this model see \citet{Zhang}.

@@

{\color{blue} MPK edit this}
\section{Results and Discussions}\label{sec:results}

@@

The same caveat is present in the MIXAE implementation. While it shows better optimal performance than the K-means method, some inconsistencies were observed, that were notably not evident from the unsupervised training-objectives of the model. These two factors currently conspire to limit its immediate applicability, and more developments are needed for this approach.
